---
title: "Homework - QMSS G5072"
author: "Anna Takacs"
date: "11/11/2019"
output: html_document
---

```{r include=FALSE}
# install.packages("httr")
library(httr)
# install.packages("jsonlite")
library(jsonlite)
library(tidyverse)
library(dplyr)
```

For the exercise, we will be using LA crime data. Here is the description of the API: https://dev.socrata.com/foundry/data.lacity.org/63jg-8b9z 

# 1. Define the base URL

__In the documentation, find the API Endpoint for csv (Hint: if there is a json at the end, replace json with csv).__
```{r}
endpoint <- "https://data.lacity.org/resource/63jg-8b9z.csv"
```

# 2. Send a simple GET request

__Using httr, send a simple GET request to the base URL. Print the status...__

```{r}
LAcrime <- GET(endpoint)
LAcrime
```

__and structure of the response (print only the first level in your ouput of str using the option max.level).__
```{r}
str(LAcrime, max.level = 1)
```

# 3. Get the content of the response

__Retrieve the content of the response. What are the dimensions of the data frame? Tabulate the areas of the observations included.__
```{r}
crime_data <- content(LAcrime)
# If I understand the question right, I tabulate the areas of the observations included.
names(crime_data)
```

# 4. Get your access token

__As you have noticed, our initial request yielded only a small part of the data. For full access (actually the API seems not to care much, but for exercise’s sake let’s do it), you will need an API key to access the data. Sign up here https://data.lacity.org/profile/app_tokens and get your token.__

__To keep your key safe, it is always better not to add it to your code. Place your token in the R environment file (as shown in class). Add the following lines (with your access info) to the .Renviron file (the usethis::edit_r_environ() function will be useful).__
__Note: if you have trouble, take a look at this more extended discussion of the .Renviron file.__
```{r}
R.home(component = "home")
usethis::edit_r_environ()
```

```{r}
queryparams <- list(
  limit = 10000,
  app_token = Sys.getenv("LACRIME_APP_TOKEN")
  )
```

```{r}
path <- "https://data.lacity.org?"
LAcrime_withquery <- GET(path, 
                         query = queryparams)
LAcrime_withquery$status_code
```
__Now, resend your query with the fields: "$limit" = "10000" and "$$app_token" = Sys.getenv("LACRIME_APP_TOKEN") as query parameters and show the status of the server response.__

# 5. Limiting your query

__Choose at least two more query field parameters from the API description [the blueish list on the bottom of the page], pass them to the query parameter list, and run the query (providing the server status again). Describe briefly in words what your query is selecting.__

```{r}
newquery <- list(
  limit = 10000,
  app_token = Sys.getenv("LACRIME_APP_TOKEN"), 
  vict_sex = F,
  area = 5
  )
LAcrime_newquery <- GET(path, 
                         query = newquery)
LAcrime_newquery$status_code
```
It selects those observations where the victim was female from the area which code is 5. 

# 6. Locations

__The dataset already provides us with good geographic information on latitude and longitude and some address information (for privacy reasons, the locations are approximate).__

__Let’s try to use the Google Maps or the OpenStreetMap reverse geocode APIs to find the addresses for some locations.__

___Open Street Maps___
__If you are uncomfortable with providing some payment information to Google, you can use the more capacity-limited OpenStreetMap API. For this small data set that will be entirely sufficient.__

__The query will need to look like this:__

https://nominatim.openstreetmap.org/reverse?lat=-37.866963&lon=144.980615

```{r}
key <- "https://nominatim.openstreetmap.org/reverse?lat=-37.866963&lon=144.980615"
geo_data <- GET(key)
geo_data$status_code
```

__Choose one of the reverse geocoding APIs. Your task is to find the address of a given location in the previous dataset.__

```{r}
crime <- content(LAcrime)
crime_df <- as.data.frame(crime)
# The variables that we want to look at are the 'lat' and 'lon' variables. And thus, I will filter for rows where those variables take on values and drop those that are NAs.
crime_df %>% 
  drop_na(lat) %>%
  drop_na(lon) %>%
  select (c(lat, lon))
```

```{r}
the_50 <- sample(nrow(crime_df), 50) # And I randomly choose 50 addresses from my dataframe. 
```


__In the loop, add a status check (using an if statement) that will warn us if something goes awry. Lastly, add a Sys.sleep pause of 0.1 seconds after each URL retrieval to make sure we use the Maps API responsibly.__

```{r}
URL <- paste("https://nominatim.openstreetmap.org/reverse?", "lat=33.9825", "&", "lon=-118.2695" , sep = "")
location <- GET(URL)
location$status_code
if(http_error(location)){
  warning("There is an error.")
} else {
  content(location)
  Sys.sleep(0.1)
}

location_content <- content(location)
location_content
# This is the output for one of the locations. 
```

