---
title: "Homework 1 - Data Mining"
author: "Anna Takacs"
date: "9/23/2019"
output:
  pdf_document: default
  html_document: default
---
# Exercise 1
*In what ways does a matrix differ from a array in R? Use R code to create one of each.*

The difference between matrices and arrays is that arrays can hold multidimensional rectangular data. Rectangular means that for each dimension, the lenth of the data need to be the same (e.g. each column has the same length). Matrices are only two dimensional items. 

The following is a three dimensional matrix: [^1]
```{r echo = FALSE}
array_3_d <- array(
  1:21, dim = c(3, 3, 3), dimnames = list(
    c("one", "two", "three"), 
    c("egy", "kettő", "három"),
    c("First", "Second", "Third")
  )
)
print(array_3_d)
```
And a two dimensional matrix:
```{r echo = FALSE}
matrix_2_d <- matrix(1:9, nrow = 3, dimnames = list(
  c("egy", "kettő", "három"),
  c("one", "two", "three")
))
print(matrix_2_d)
```
[^1]: Columns are named in Hungarian. 

# Exercise 2
Simplify and make the following code consistent with the Hadley Wickham R style guide.
```{r}
a <- c(5, -2, 3, -4, 1, 2) 
b <- a*-1
b [b > 0]

num = 1
mycondition <- round(runif(1, 0, 1)) 
if( mycondition ) {num <- num + 1
print(paste("num =", num, sep = " "))} else {print("false")}
```
Firstly, create an integer sequence from 1 to 50. To see if a number if even, use the module function. And lastly, subset the sequence of integers depending on whether they are divisible by 2.
```{r}
y <- 1:50
even <- y %% 2 == 0
y = y [even]
```

The following code creates a matrix and the mean of the values in the matrix divided by the number of elements it has. 
```{r} 
x <- matrix(c(23, 34, 35, 6, 87, 39, 21, 14, 99), nrow = 3) 
df <- as.data.frame(x)
names(df) <- c("percentage_score_on_reading_test",
 "percentage.score.on.math.test",
  "percentage-score-on-writing-test")
my_mean <- function(x) {sum(x) / length(x)}
```

# Exercise 3
*Look at the spredsheet-like representation of the data and write an R code that extracts the data for Ohio on the variables 'population' and 'frost' in three different ways.*
```{r echo = FALSE, include = FALSE}
str(state.x77)
```

1st way of extracting: 
```{r}
state.x77["Ohio", c("Population", "Frost")]
```

2nd way of extracting:
```{r}
state.x77["Ohio", -c(2:6, 8)]
```

3rd way of subtracting information: 

I transform the tibble into a data frame. 
In this way, I can tell R STudio to extract the values from the 35th row and 1st and 7th columns. 
```{r}
class(as.data.frame(state.x77))
state.x77[35, c(1, 7)]
```

# Exercise 4 
```{r include = FALSE, echo = FALSE}
help(airquality)
```
*Replace the Wind variable with windspeed measured in kilometers per hour.*
```{r}
# I multiply the value of the wind column in the dataset by 1.609344 so that the values are expressed in km/h instead of mph.
airquality$Wind <- airquality$Wind * 1.609344
# To be able to distingues between the new and old variables, I rename the new km/h variable to windspeed
names(airquality) <- c("Ozone", "Solar.R", "Windspeed", "Temp", "Month", "Day")
print(head(airquality))
```

# Exercise 5
I brings the data.frame called turnout into R’s memory and observe its values. 
```{r echo = FALSE }
data("turnout", package = "qss")
help("turnout", package = "qss")
print(head(turnout))
```

To calculate the turnout rate based on the voting age population, I add the voting age population and the number of eligible oversees voters and multiply it with the estimated turnout rate to get a value for the turnout rate. I repeat the same process with the voting eligible population. I print out the values for both for the corresponding years. 
```{r echo = FALSE}
turnoutrate_age <- (turnout$VAP + turnout$overseas) * turnout$ANES
print(turnoutrate_age)
turnoutrate_eligible <- (turnout$VEP + turnout$overseas) * turnout$ANES
print(turnoutrate_eligible)
```
I observe that the turnout rate is lower when I use the voting eligible value, which is intuitive coming from the fact that not everyone is eligible to vote who are above the voting age, but everyone who is eligible to who must be above the voting age. 

> This concludes the solutions for homework 1 for the Data Mining class. 

